Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\

### 2025-03-10 main-disk-model.tex
further cleaned main-disk-model

In the {\bf offline setting}, the goal is to find a {\em schedule} of the DAG~$G$ for the $k$ processors,
so that the makespan (total execution time) is minimized. Formally, a schedule contains:
\begin{itemize}
\item for each task, its processor allocation and starting time, 
as well as the amount of data offloaded for the execution of the task; 
\item for each read and write operation to disk, the starting time of the operation; 
\item for each data and processor, the time intervals during which the file is available
on the local memory of the processor (once it has been generated or read, and before 
it is used or evicted). 
\end{itemize}

@@

%\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
%I'll probably add a few sentences here...}
% while respecting memory constraints. \skug{do we still formulate it like that?}
    %If a processor runs out of memory to execute
%a task mapped on it, the schedule is said to be {\em invalid}.

@@

Note that the offline problem is already NP-hard even in the homogeneous case and 
without memory constraints (i.e., no need to offload data onto disk), and even for
a graph without precedence constraints (independent tasks). 
%because of the DAG structure of the application. 
Hence, we focus on the design of efficient scheduling heuristics. 
%\hmey{Mention NP-hardness due to being more general than NP-hard problem?}

%    \paragraph{Workflow-related changes}
%
%    \begin{itemize}
%        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
%        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
%
%    \end{itemize}
%
%    The following changes are not a part of this article's scope:
%
%    \begin{itemize}
%        \item The workflow structure changes: edges or tasks come in or leave.
%    \end{itemize}
%
%    \paragraph{Execution environment-related changes }
%
%
%    \begin{itemize}
%        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
%        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
%
%    \end{itemize}
%
%    The following changes are not a part of this article's scope:
%
%    \begin{itemize}
%        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
%    \end{itemize}
%
%    \subsection{Time of changes }
%
%    We consider discrete time in seconds.
%    The time point(s) at which the changes happen is unambiguously defined.
%
%    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
%    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
%    The start time of the source task in the workflow is zero.
%    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
%
%    \subsection{Changes and knowledge horizon - important questions TBA}
%
%    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
%
%    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
%    However, how do we model the following:
%    \begin{itemize}
%        \item Do we know the new weights of currently running tasks and tasks that have not yet started? This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
%        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
%        \item How many times of change do we model - one per workflow run, or multiple?
%        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
%    \end{itemize}

\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
\label{sec:heuristics}
\subsection{\skug{New scheduling approach - general description}}

@@

\subsection{Data structures }
    To optimally represent the timeline of the execution of a workflow, we represent the execution as a series of events
    with attached unique timestamps.
    The events are kept in the priority queue ordered by timestamp.
    Events are created when other events are triggered and are being inserted into the queue.
    When their timestamp is the smallest, they are then triggered, possibly creating new events.
    The execution starts with placing the event of starting the starting node into the queue with the timestamp 0.
    The execution ends when there are no events in the queue.

%    The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
%    but favor processor reuse as is done in HEFT.
%    Furthermore, this would allow us to handle variability on the fly, by updating
%    the bottom levels if some parameters vary, and computing the schedule
%    only for the near future...

%In order to be able to easily adapt to variability of task parameters, 
We design 
variants of HEFT that account for memory usage and aim at minimizing the makespan.
First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic that does not account for the memory
(and hence, may return invalid schedules that will not be able to run on the platform
by running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
heuristics, including eviction strategies to move some data in communication buffers
in case there is not enough memory available on some processors.

@@

>>>>

In the {\bf offline setting}, the goal is to find a {\em schedule} of the DAG~$G$ for the $k$ processors,
so that the makespan (total execution time) is minimized. Formally, a schedule contains:
\begin{itemize}
\item for each task, its processor allocation and starting time, 
as well as the amount of data offloaded for the execution of the task; 
\item for each read and write operation to disk, the starting time of the operation; 
\item for each data and processor, the time intervals during which the file is available
on the local memory of the processor (once it has been generated or read, and before 
it is used or evicted). \skug{not sure if we need this interval?}
\end{itemize}

@@

@@

Note that the offline problem is already NP-hard even in the homogeneous case and 
without memory constraints (i.e., no need to offload data onto disk), and even for
a graph without precedence constraints (independent tasks). 
%because of the DAG structure of the application. 
Hence, we focus on the design of efficient scheduling heuristics.

\section{Scheduling heuristics}
\label{sec:heuristics}

\subsection{Offline setting}

@@

We design 
variants of HEFT that account for memory usage and aim at minimizing the makespan.
First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic that does not account for the memory
(and hence, may return invalid schedules that will not be able to run on the platform
by running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
heuristics, including eviction strategies to move some data in communication buffers
in case there is not enough memory available on some processors.

@@

    \subsection{Online setting }
    To optimally represent the timeline of the execution of a workflow, we represent the execution as a series of events
    with attached unique timestamps.
    The events are kept in the priority queue ordered by timestamp.
    Events are created when other events are triggered and are being inserted into the queue.
    When their timestamp is the smallest, they are then triggered, possibly creating new events.
    The execution starts with placing the event of starting the starting node into the queue with the timestamp 0.
    The execution ends when there are no events in the queue.
